---
title: "StatsHW3"
output:
  pdf_document: default
  html_document: default
date: "2025-02-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
_Chapter 3_

## Question 23

```{r}
library(dplyr)
library(Sleuth3)

skin_cancer <- ex0323
#a) There may be a spatial correlation, however the sample size should resolve this. 

par(mfrow = c(1, 2))
high_sunspot <- subset(ex0323, SunspotActivity == "High")
low_sunspot <- subset(ex0323, SunspotActivity == "Low")

plot(high_sunspot$Year, high_sunspot$SkinCancerRate, 
     main = "High Sunspot Activity (2 years prior)")
plot(low_sunspot$Year, low_sunspot$SkinCancerRate, 
     main = "Low Sunspot Activity (2 years prior)")
par(mfrow = c(1, 1))

```
## Question 25

```{r}
orange <- case0302

without_outliers <- orange |>
  filter(Dioxin < 20) 
boxplot(Dioxin ~ Veteran, data = without_outliers)

```

## Question 29

```{r}
rain <- case0301
boxplot(Rainfall ~ Treatment, data=rain)
unseed <- rain |> filter(Treatment == "Unseeded")
boxplot (unseed$Rainfall +100, unseed$Rainfall)
boxplot (unseed$Rainfall +200, unseed$Rainfall)
boxplot (unseed$Rainfall +300, unseed$Rainfall)
boxplot (unseed$Rainfall +400, unseed$Rainfall)
```

_Part b_ 
```{r}
boxplot(unseed$Rainfall*2, unseed$Rainfall)
boxplot(unseed$Rainfall*3, unseed$Rainfall)
boxplot(unseed$Rainfall*4, unseed$Rainfall)
boxplot(unseed$Rainfall*5, unseed$Rainfall)

```

_Part c_

The multiplicative plots more closely match the actual data. 

## Question 33

```{r}
mammal_brain <- ex0333

boxplot(BrainSize ~ LitterSize, data = mammal_brain)
t.test(BrainSize ~ LitterSize, data = mammal_brain)
```

The difference between brain sizes depending on mammal litter size is statistically significant with a 0.016 p-value, with the Large group mean being 10.9, a mean difference of 4.1 from the Small group.

## Problem A - Didgeridoo

_Part a_ 

Because all four of the measurements are negative, which indicates the sleep apnoea related symptoms improved with the playing of a digeridoo, compared to the control group. 

_Part b_

The quality of sleep had a pvalue of 0.27, which is not independently statistically significant. 

_Part c_

The most significant measure of improvement was the decrease in partner reported sleep disturbances. 

## Problem B - Vaccine

_I wasn't sure exactly how you wanted us to solve this, we've talked about the concept in class, but I might have missed on the exact part/ formula of how to solve it. I used the internet for some help._

```{r}
delta <- 1.0     
sd <- 0.73
alpha <- 0.05     
power <- 0.99     

sample_size <- power.t.test(delta = delta, sd = sd, sig.level = alpha, power = power, type = "two.sample")$n
ceiling(sample_size)

```

This result of 21 students per sample size means a total sample size of 42 students. 



