---
title: "statshw4"
author: "Katrina Churchill"
date: "2025-02-24"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 4
_conceptual exercises 2,3,9_
Hand in 19, 26, 27, 28, 29, 31, 32

## Question 19 

```{r}
library(dplyr)
library(Sleuth3)

#Part a
t.test(Humerus ~ Status, data=ex0221)

#Part b
#t.tests are based on a normal data spread assumption. 

#Part c
#Yes, a continunity correction would be needed for the Z-statistic. 

#Part d 
wilcox.test(Humerus ~ Status, data=ex0221)

min(ex0221$Humerus)
birds_wo_smallest <- ex0221 |> filter(Humerus > 0.659)

t.test(Humerus ~ Status, data=birds_wo_smallest)
wilcox.test(Humerus ~ Status, data=birds_wo_smallest)

#At first, the t.test pvalue = 0.09236 and the wilcox.test = 0.1718
#After removing the smallest value, the t.test pvalue =  0.176 and 
#the wilcox.test = 0.2588

#part e
#Wilcox tests are resiliant to outliers, while t.test are not, shown by 
#removing the outlier and the t.test pvalue almost doubling.
```

## Question 26

```{r}
#Part a 
differences <- case0202$Unaffected - case0202$Affected
hist(differences, breaks = 20)
t.test(case0202$Unaffected, case0202$Affected, paired = TRUE)
# There appears to be a right skew of the differences. 

#Part B
# Compute the log-transformed hippocampus volumes
log_affected <- log(case0202$Affected)
log_unaffected <- log(case0202$Unaffected)

log_differences <- log_unaffected - log_affected
hist(log_differences, breaks = 20)
#The data spread does not appear to be more symmetric

#Part c
t.test(log_unaffected, log_affected, paired = TRUE)
#The pvalue after a log transformation is 0.006463, compared to a 0.006062 
#before transforming the data, a very small difference. 

#Part d
mean_log_diff <- mean(log_differences)
se_log_diff <- sd(log_differences) / sqrt(length(log_differences))

t_critical <- qt(0.975, df = length(log_differences) - 1) 
ci_lower_log <- mean_log_diff - t_critical * se_log_diff
ci_upper_log <- mean_log_diff + t_critical * se_log_diff

median_ratio_estimate <- exp(mean_log_diff)
ci_lower_ratio <- exp(ci_lower_log)
ci_upper_ratio <- exp(ci_upper_log)

cat("Mean Log Difference:", mean_log_diff, "\n")
cat("95% CI for Log Difference: (", ci_lower_log, ",", ci_upper_log, ")\n")
cat("Estimated Median Ratio:", median_ratio_estimate, "\n")
cat("95% CI for Median Ratio: (", ci_lower_ratio, ",", ci_upper_ratio, ")\n")

```

## Question 27

```{r}
wilcox.test(log_affected, log_unaffected, paired = TRUE)
wilcox.test(case0202$Affected, case0202$Unaffected, paired = TRUE)

#The pvalue of the two tests above match because wilcox.test doesn't change with
# a log transformation, because the ranking is the same. 

#It doesn't seem apparent that the assumptions of the rank test are any more 
#appropiate on these scales

```

## Question 28

```{r}
#Part a
differences_plant <- ex0428$Cross - ex0428$Self
hist(differences_plant)

#Part b
t.test(ex0428$Self, ex0428$Cross, paired = TRUE)

#Part c
t.test(ex0428$Cross, ex0428$Self, paired = TRUE, conf.level = 0.95)
#I believe this returns the correct confidence interval, I did a different
#method for an earlier question. 

#Part D
#The histogram of part a is suggesting quite a large discrepancy between 
#cross and self fertilized plants, while the pvalues of our tests are not or 
#barely reaching significance. 

#Part e 
wilcox.test(ex0428$Self, ex0428$Cross, paired = TRUE)
```

## Question 29

```{r}
#Part a
salvage <- ex0429
ex0429$Action <- as.integer(ex0429$Action)
wilcox.test(ex0429$Action, ex0429$PercentLost)

wilcox.test(ex0429$Action, ex0429$PercentLost, conf.int = TRUE, conf.level = .95)

#Part b
t.test(PercentLost ~ Action, data = ex0429, conf.level = 0.95)
```
Between part a and part b, the pvalues were 0.0002534 and 0.008849 respectively. This indicates that the t.test might be missing a problem with the data, possibly an outlier that is skewing significance. Both conclude a strong confidence level. 

##Question 30

```{r}
sun <- ex0430
#because I do not see any obvious outliers, I will be using a t.test analysis. 
sunscreen <- ex0430$Sunscreen
before <- ex0430$PreTreatment

t.test(before, sunscreen, paired = TRUE, conf.level = .95)
```
Personally, there are so many things I will consider when I am tolerating the sun that aren't sunscreen, such as hydration, clothing coverage, direct or indirect sun exposure, outside/ inside, how humid it is, how mch time between the first test, applying, the sunscreen, and going back into the sun, etc. The genetic makeup/ pre-existing conditions could have a big impact too, as someone with a heart condition might not stay in the sun as long the second round. I'm fairly certain women have higher pain tolerances, but worse body temperature regulation. 
